---
title: "nbaR API client: concepts and object model"
date: "`r format(Sys.time(), '%b %d, %Y')`"
output:
  html_document:
    highlight: tango
    code_folding: show
    theme: sandstone
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false

vignette: >
  %\VignetteIndexEntry{nbaR API client: concepts and object model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---



## Main data types and classes
The data served by the NBA consists of four main data types:
  
  * Specimen
  * Taxon
  * Multimedia
  * Geo
  
Additionally, the data type *Metadata* stores miscellaneous
information about NBA settings. Each of the data types is modelled
as an `R6` class and therefore has its own members such as fields
and methods. Documentation about a specific class can be retrieved
in the standard manner, e.g. `?Specimen`
Each class of the data model can be instantiated and has a `toJSONString` and `toList` method returning
the object's JSON representation and the object's data as a list datatype, respectively.

```{r}
# load nbaR
library(nbaR) 

# instanciate specimen object
spec <- Specimen$new()
```
```r
# represent data as JSON or list
spec$toJSONString()
spec$toList()
```

## API Client classes
The interaction with the API is accomplished by the API client classes:

* SpecimenClient
* TaxonClient
* MultimediaClient
* GeoClient
* MetadataClient


```{r}

# initialize client
client <- SpecimenClient$new()
```

The client class is by default initialized to connect to the base URL http://api.biodiversitydata.nl/v2.
For testing purposes, this can be set to a different URL, see `?SpecimenClient` for details.

## Queries

### Concept

With the SpecimentClient created above, the `query` endpoint for specimens can now be reached via the `query` function.
Query parameters can be specified as a list with named parameters.
To query for instance for specimen records that have the type status holotype and a female sex,
one can pass a named list as the `queryParams` parameter to the `query` function:


```{r}
# specify two query conditions
l <- list(identifications.typeStatus="holotype", sex="female")

# run query
res <- client$query(queryParams=l)

```

The `query` function then returns an object of class `Response`, which, in turn, has a field `content`
of class `QueryResult`. From the `QueryResult`, the single result items can be accessed as follows:

```{r}
# get first element of results
spec <- res$content$resultSet[[1]]$item

# object should be of type Specimen
class(spec)

```

Note that by default, both conditions are connected by a logical `AND`. 
The `queryParams` passed as a list thus correspond to 
basic [*human readable* queries](http://docs.biodiversitydata.nl/en/latest/quickstart/#human-readable).
For more advanced queries, containing different logical operators or nested sub-queries, 
the user can specify the query in a `QuerySpec` object.


# Advanced queries

## Using the QuerySpec object
Advanced queries with different operators than `AND` or nested query conditions 
can not be accomplished by simply passing the query parameters as a list. 
Instead, a query is modeled as a `QuerySpec` object which
captures the relationships between multiple query terms. Please also refer to the 
[NBA QuerySpec documentation](http://docs.biodiversitydata.nl/en/latest/advanced-queries/#queryspec) for more information.

A `QuerySpec` object usually consists of one or more `QueryCondition` objects, specifying 
query terms. A `QueryCondition` object usually contains the fields `field`, `operator`, and `value `(see also `?QueryCondition`).
These fields can be specified in the constructor. If, for example, we want to query for records with a unitID equal to
*L.4304195*, a `QueryCondition` would look as follows:

```{r}
# make specimen client instance
client <- SpecimenClient$new()

# specify search in QueryCondition object
qc <-
  QueryCondition$new(field = "unitID",
                     operator = "EQUALS",
                     value = "L.4304195")
```

Note: The function `get_field_info` on a certain field lists which operators 
are allowed for that field e.g. for the field `unitID` : 

```{r}
client$get_field_info()$content$unitID$allowedOperators
```

Now, a `QuerySpec` object can be assembled with the `QueryCondition`(s) passed as a list:

```{r}
# build QuerySpec using above conditions
qs <- QuerySpec$new(conditions=list(qc))

# do the query
res <- client$query(querySpec=qs)
```

Below, we show an example of how to nest multiple query conditions. The query conditions below define to query for 
specimens of sex ** female* and family *Equidae* and of rank *Species*. 

```{r}
# specify multiple conditions
q1 <- QueryCondition$new(field = "sex",
                         operator = "EQUALS",
                         value = "female")
q2 <-
  QueryCondition$new(field = "identifications.defaultClassification.family",
                     operator = "EQUALS",
                     value = "Equidae")
q3 <- QueryCondition$new(field = "identifications.taxonRank",
                         operator = "EQUALS",
                         value = "species")
```

Extending the constraint to also include specimens of rank *Subspecies*,  
we can combine the latter condition with an additional one using the method `or`:

```{r}
# logical conjunction using 'or' method
q3$or <- list(QueryCondition$new(field = "identifications.taxonRank",
                                 operator="EQUALS",
                                 value="subspecies"))

# build QuerySpec from QueryConditions
qs <- QuerySpec$new(conditions=list(q1, q2, q3))

# call API
res <-client$query(querySpec=qs)
```

### Size of the query result set
By default, the NBA returns the first 10 hits for a given query. 
In, for instance, a query without parameters has many hits

```{r}
res <- client$query()
res$content$totalSize
```

but only the first 10 are returned in the `resultSet`: 

```{r}
length(res$content$resultSet)
```

In order to increase the size of a `resultSet`, a `size` parameter can be
passed to the constructor of a `QuerySpec` object. Below, we will get
the first 1000 records of the query above:

```{r}
qs <- QuerySpec$new(size=1000)
res <- client$query(querySpec=qs)
length(res$content$resultSet)
```

