---
title: "nbaR: R access to the Netherlands Biodiversity API"
date: "`r format(Sys.time(), '%b %d, %Y')`"
output:
  html_document:
    highlight: tango
    code_folding: show
    theme: sandstone
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false

vignette: >
  %\VignetteIndexEntry{nbaR: R access to the Netherlands Biodiversity API}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Background
  
## The Netherlands Biodiversity API

The Netherlands Biodiversity API (NBA) facilitates access to the
Natural History Collection at Naturalis Biodiversity Center. Next to
museum specimen records and metadata, access to taxonomic
classification and nomenclature, to geographical information, and to
multimedia files is provided. By using the powerful Elasticsearch
engine, the NBA facilitates searching for collection- and biodiversity
data in near real-time. Furthermore, by incorporating information from
taxonomic databases, taxonomic name resolution can be accomplished
with the NBA. Persistent Uniform Resource Identifiers (PURLs) ensure
that each specimen accessible via the NBA is represented by a citeable
unambiguous web reference. Access to our data is provided via a
RESTful interface and several clients such as the
[BioPortal](http://bioportal.naturalis.nl/), a web application for
browsing biodiversity data that is served by the NBA.  For more
information about the NBA, please see our
[detailed documentation](http://docs.biodiversitydata.nl/).
  
## R access

The R programming language is established as a common tool in
scientific research, with growing adoption by researchers in
biodiversity research.  Hence, to ease the access to the NBA for
researchers, we developed this R client.
  
# Full client vs wrapper functions

nbaR aims to be a *full* client of the NBA API, meaning that it
implements all endpoints and the entire NBA object model.  The client
thus facilitates all API queries possible.  Complex objects returned
by the API, such as `Specimen` or `Taxon` objects are implemented as
`R6` classes. This includes also objects used for querying
(`QuerySpec` and `QueryCondition`, respectively, see also
[here](apiclient.html)).

For for many queries, the full functionality of the NBA won't be
required. The package therefore offers a **wrapper function** for each
endpoint that does not use `R6` classes but common R data structes,
such as `list` and `data.frame`. However, querying capabilities are
limited for these wrappers. Below we will show how to set up some
simple queries using the wrapper functions.

# Quick start: querying the NBA using wrapper functions

The data in the NBA consists of four main data types (see
[NBA docs](http://docs.biodiversitydata.nl/)):
  
  * Specimen
  * Taxon
  * Multimedia
  * Geo

Wrapper functions start with the data type (lower-case letter) and an
underscore (`specimen_*`, `taxon_*`) etc. There is a wrapper function
for each endpoint (see
[here](https://docs.biodiversitydata.nl/endpoints-reference/) for all
endpoints); camelCase naming is replaced by snake_case. The NBA
endpoint `getDistinctValues` for specimen data, for instance, is
called by the function `specimen_get_distinct_values`.

Specimen services provide the interface to the Naturalis collection
and to species occurrences (see
[here](https://docs.biodiversitydata.nl/en/latest/doc-spec-services/specimen/)),
wheras Taxon services provide data from taxonomic checklists (see
[here](https://docs.biodiversitydata.nl/en/latest/doc-spec-services/taxon/)).
Multimedia services give access to photos, videos and sound data (see
[here](https://docs.biodiversitydata.nl/en/latest/doc-spec-services/multimedia/));
Geo services store polygon data for geographical regions and nature
reserves (see
[here](https://docs.biodiversitydata.nl/en/latest/doc-spec-services/geo/)).

## Querying specimen records

Suppose we want to look up specimens of the genus *Mola* (sunfish). 
To find out what field of the NBA we could query, we can use the 
function `specimen_get_paths()` (see `?specimen_get_paths` for 
documentation).

```{r}
library('nbaR')

all_paths <- specimen_get_paths()
head(all_paths)
```

Note that paths of nested objects are seperated via a `.`.  To search
for a specific genus, we can query the field
`identifications.scientificName.genusOrMonomial`.  The
`specimen_query` method lets us query for a specific field, where the
query parameters are given as a named list (a named vector also
works!):

```{r}
queryParams <- list("identifications.scientificName.genusOrMonomial" =
                        "Mola")
sp_data <- specimen_query(queryParams)

## how many specimens are found?
nrow(sp_data)

## which fields are available?
colnames(sp_data)
```

Return type can either be `list` or `data.frame` (the default).  Note
that nested structures in the data frame are represented as list
columns (for instance the field `associatedMultiMediaUris`). which
lists, if given, all links to multimedia resources for the specimens:

```{r}
sp_data$associatedMultiMediaUris
```


## Querying taxon records

Taxonomic information can be retrieved using the **taxon_** functions.
Taxon records come from two sources, the *Dutch species register
(Nederlands Soortregister, NSR)* and the *Catalogue of Life (COL)*.

To see how many records are from each source, we can query for all
distinct values (and counts) for a specific field (see
`taxon_get_paths`) for all fields in the `taxon` data:

```{r}
taxon_get_distinct_values("sourceSystem.name")

## alternatively, show for sourceSystem.code
taxon_get_distinct_values('sourceSystem.code')
```

To query, for instance all the species listed in the Catalogue of life
for the genus *Mola*, we can use the wrapper function `taxon_query`:

```{r}
## specify query parameters
queryParams <- list("sourceSystem.code"="COL",
                    "defaultClassification.genus"="Mola")

## do the query
tax_data <- taxon_query(queryParams)

## access nested field 'accepted Name' -> 'specificEpithet'
tax_data$acceptedName$specificEpithet
```

Let's see if we can find vernacular (common) names for the
species **Mola ramsayi**:

```{r}
tax_data$vernacularNames[[3]]
```

## Geo queries

The Geo data type in the NBA holds polygon data for countries, Dutch
municipalities etc, and Dutch nature reserves. For more information
please refer to the
[API documentation](https://docs.biodiversitydata.nl/en/latest/doc-spec-services/geo/).
To retreive e.g. a polygon, encoded in the
[geoJSON](http://geojson.org/) format for a country, we can query as
follows:


```{r}
geo_json <- geo_get_geo_json_for_locality('Nigeria')
```

## Multimedia queries

Multimedia items accessible via the NBA include items captured from
physical specimens (e.g. photos and videos) but also from human
observations (e.g. recordings of bird sounds).

As an example, we will retrieve records that represent sounds that
were recorded in the country *Cape Verde*.  The sound data accessible
via the NBA is stored in the
[Xeno-Canto database](https://www.xeno-canto.org), hosted at the
Naturalis Biodiversity Center.  The field `sourceSystem.code` for
these records is `XC`; the country of occurrence is stored in the
field `gatheringEvents.country`.

```{r}
queryParams <- list("sourceSystem.code"="XC",
                    "gatheringEvents.country"="Cape Verde")

mm_data <- multimedia_query(queryParams)

## Access link to Xeno-Canto database for each record:
mm_data$recordURI
```

## Limitations of wrapper functions

It is important to note that querying power is limited using the
wrapper functions. They relate to basic, human readable NBA queries
([see here](https://docs.biodiversitydata.nl/en/latest/quickstart/)).

 * **Size of result set:** As by NBA default, wrapper functions 
   only return the first 10 hits of a query. 
 * **Operators:** Only full matches (operator *EQUALS*) are 
   considered in wrapper query functions. Partial matching is
   only available in the full API client
 * **Locical conjunctions:** If multiple query conditions are
   given, wrapper functions only allow a simple *AND* conjunction.
   For more complex logical query constructs including *OR* 
   operators or negations, the full API client must be used.

The wrappers are thus designed for easy access for simple queries.  In
many situations it might be necessary to use the full API client which
offers (almost) the entire functionality of the NBA API. Detailed
documentation for the full client can be found [here](apiclient.html).

 
