---
title: "nbaR: R access to the Netherlands Biodiversity API"
date: "`r format(Sys.time(), '%b %d, %Y')`"
output: 
  html_document:
    highlight: tango
    code_folding: show
    theme: sandstone
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      
vignette: >
%\VignetteIndexEntry{nbaR: R access to the Netherlands Biodiversity API}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  ---
  
  # Background
  
  ## The Netherlands Biodiversity API
  The Netherlands Biodiversity API
  (NBA) facilitates access to the Natural History Collection at 
  Naturalis Biodiversity Center. Next to museum specimen records
  and metadata, access to taxonomic classification and nomenclature,
  to geographical information, and to multimedia files is provided. By
  using the powerful Elasticsearch engine, the NBA facilitates searching
  for collection- and biodiversity data in near real-time. Furthermore,
  by incorporating information from taxonomic databases, taxonomic name
  resolution can be accomplished with the NBA. Persistent Uniform
  Resource Identifiers (PURLs) ensure that each specimen accessible via
  the NBA is represented by a citeable unambiguous web reference. Access
  to our data is provided via a RESTful interface and several clients such
  as the [BioPortal](http://bioportal.naturalis.nl/), a web application
  for browsing biodiversity data that is served by the NBA.  For more
  information about the NBA, please see our
  [detailed documentation](http://docs.biodiversitydata.nl/).
  
  ## R access
  The R programming language is established as a common tool in scientific 
  research, with growing adoption by researchers in biodiversity research.
  Hence, to ease the access to the NBA for researchers, we developed this R 
  client.
  
  # Getting started
  
  ## Main data types and classes
  The data served by the NBA consists of four main data types:
  
  * Specimen
  * Taxon
  * Multimedia
  * Geo
  
  Additionally, the data type *Metadata* stores miscellaneous
  information about NBA settings. Each of the data types is modelled
  as an `R6` class and therefore has its own members such as fields
  and methods. Documentation about a specific class can be retrieved
  in the standard manner, e.g. `?Specimen`
  Each class of the data model can be instantiated and has a `toJSONString` and `toList` method returning
  the object's JSON representation and the object's data as a list datatype, respectively.
  
  ```{r}
# load nbaR
library(nbaR) 

# instanciate specimen object
spec <- Specimen$new()
```
```r
# represent data as JSON or list
spec$toJSONString()
spec$toList()
```

## API Client classes
The interaction with the API is accomplished by the API client classes:

* SpecimenClient
* TaxonClient
* MultimediaClient
* GeoClient
* MetadataClient


```{r}

# initialize client
client <- SpecimenClient$new()
```

The client class is by default initialized to connect to the base URL http://api.biodiversitydata.nl/v2.
For testing purposes, this can be set to a different URL, see `?SpecimenClient` for details.

## Queries

### Concept

With the SpecimentClient created above, the `query` endpoint for specimens can now be reached via the `query` function.

```{r}
# retrieve results of empty query.
res <- client$query()
```

The `query` function then returns an object of class `Response`, which, in turn, has a field `content`
of class `QueryResult`. From the `QueryResult`, the single result items can be accessed as follows:

```{r}
# get first element of results
spec <- res$content$resultSet[[1]]$item

# object should be of type Specimen
class(spec)

```

### Passing query parameters as list
Query parameters can be specified as a list with named parameters.
To query for instance for specimen records that have the type status holotype and a female sex,
one can pass a named list as the `queryParams` parameter to the `query` function:

```{r}
# specify two query conditions
l <- list(identifications.typeStatus="holotype", sex="female")

# run query
res <- client$query(queryParams=l)
```

Note that by default, both conditions are connected by a logical `AND`. 
The `queryParams` passed as a list thus correspond to 
basic [*human readable* queries](http://docs.biodiversitydata.nl/en/latest/quickstart/#human-readable).
For more advanced queries, containing different logical operators or nested sub-queries, 
the user can specify the query in a `QuerySpec` object.

# Advanced queries

## Using the QuerySpec object
Advanced queries with different operators than `AND` or nested query conditions 
can not be accomplished by simply passing the query parameters as a list. 
Instead, a query is modeled as a `QuerySpec` object which
captures the relationships between multiple query terms. Please also refer to the 
[NBA QuerySpec documentation](http://docs.biodiversitydata.nl/en/latest/advanced-queries/#queryspec) for more information.

A `QuerySpec` object usually consists of one or more `QueryCondition` objects, specifying 
query terms. A `QueryCondition` object usually contains the fields `field`, `operator`, and `value `(see also `?QueryCondition`).
These fields can be specified in the constructor. If, for example, we want to query for records with a unitID equal to
*L.4304195*, a `QueryCondition` would look as follows:

```{r}
# specify search in QueryCondition object
qc <-
  QueryCondition$new(field = "unitID",
                     operator = "EQUALS",
                     value = "L.4304195")
```

Note: The function `getFieldInfo` on a certain field lists which operators 
are allowed for that field (e.g. ```sc$get_field_info()$content$unitID$allowedOperators```).

Now, a `QuerySpec` object can be assembled with the `QueryCondition`(s) passed as a list:

```{r}
# build QuerySpec using above conditions
qs <- QuerySpec$new(conditions=list(qc))

# do the query
res <- client$query(querySpec=qs)
```

Below, we show an example of how to nest multiple query conditions. The query conditions below define to query for 
specimens of sex ** female* and family *Equidae* and of rank *Species*. 

```{r}
# specify multiple conditions
q1 <- QueryCondition$new(field = "sex",
                         operator = "EQUALS",
                         value = "female")
q2 <-
  QueryCondition$new(field = "identifications.defaultClassification.family",
                     operator = "EQUALS",
                     value = "Equidae")
q3 <- QueryCondition$new(field = "identifications.taxonRank",
                         operator = "EQUALS",
                         value = "species")
```

Extending the constraint to also include specimens of rank *Subspecies*,  
we can combine the latter condition with an additional one using the method `or`:

```{r}
# logical conjunction using 'or' method
q3$or <- list(QueryCondition$new(field = "identifications.taxonRank",
                                 operator="EQUALS",
                                 value="subspecies"))

# build QuerySpec from QueryConditions
qs <- QuerySpec$new(conditions=list(q1, q2, q3))

# call API
res <-client$query(querySpec=qs)
```

### Size of the query result set
By default, the NBA returns the first 10 hits for a given query. 
In, for instance, a query without parameters has many hits

```{r}
res <- client$query()
res$content$totalSize
```

but only the first 10 are returned in the `resultSet`: 

```{r}
length(res$content$resultSet)
```

In order to increase the size of a `resultSet`, a `size` parameter can be
passed to the constructor of a `QuerySpec` object. Below, we will get
the first 1000 records of the query above:

```{r}
qs <- QuerySpec$new(size=1000)
res <- client$query(querySpec=qs)
length(res$content$resultSet)
```

# Services examples

## Specimen occurrence services {.tabset}
Specimen records constitute the core of
data served by the NBA. Museum specimens can represent a whole variety
of different objects such as plants, animals or single parts thereof,
DNA samples, fossils, rocks or meteorites. For detailed information of the
data model, please refer to the [official documentation in the NBA](http://docs.biodiversitydata.nl/en/latest/doc-spec-services/specimen/).

All specimen occurrence services are accessible using the methods within the `SpecimenClient` class. 
For a list of available endpoints, please refer to the class documentation (`?Specimen`). Below, 
we will give details about all services, grouped by category.

### Query services
#### query
Querying for specimens is accomplished with the `query` method in the class `SpecimenClient`. For simple queries, 
query parameters of type `list` can be passed via the parameter `queryParams`, for example we 
can query specimens of the Family *Ebenaceae* that were collected in Europe:

```{r}
# instantiate specimen client
sc <- SpecimenClient$new()

# specify query params in named list
qp <-
  list(
    identifications.defaultClassification.family = "Ebenaceae",
    gatheringEvent.continent = "Europe"
  )

# query
res <- sc$query(queryParams = qp)
```

If we now want to know all the countries that the specimens were collected in, we can access the `Specimen` objects in ```res$content$resultSet```
as follows:

```{r}
sapply(res$content$resultSet, function(x)x$item$gatheringEvent$country)
```

Note that passing query parameters as a named list only allows for limited queries; the logical conjunction between 
parameters is for example always `AND`. More complex queries can be accomplished using the `QuerySpec` object:

```{r}
# get all specimens with genus name starting with 'Hydro'
qc <-
  QueryCondition$new(field = "identifications.defaultClassification.genus",
                     operator = "STARTS_WITH",
                     value = "Hydro")
qs <- QuerySpec$new(conditions = list(qc))
res <- sc$query(qs)
```

#### download_query
The `query` function is limited to retrieve 50000 specimen at once (this is determined
in the parameter `index.max_result_window`, the value is retrievable using the `getSettings` method
in the metadata section). In order to provide access for a larger amount of data, 
the `query_download` takes the same arguments as `query`, but download the data as a gzip stream under the hood. 
Unlike `query`, `query_download` returns a list of specimen objects instead of a `ResultSet`.
Example:

```{r eval=FALSE}
## get the first 100000 specimen objects (not possible with query method)
res <- sc$download_query(QuerySpec$new(size=100000))
```

### Data access services
Several access methods offer the convenient retrieval of specimens matching a certain 
identifier or being part of a certain collection. Below we give examples of how to use the 
currently implemented data access services for specimen records:

#### Access to collections
Some of the specimens available via the NBA are categorised thematically into special collections, such as the Siebold-, Dubois- or Jongmans collection.
The function `get_named_collection` lists all available special collections and the identifiers of the specimens within a collection can be
queries with `get_ids_in_collection`

```{r}
sc$get_named_collections()$content
sc$get_ids_in_collection("siebold")$content
```

#### count
For any given query (with `QuerySpec` or not), returns the count of matches instead of specimen objects:

```{r}
# Example with QuerySpec:
# how many specimens are there in the 'Botany' collection?
qc <- QueryCondition$new(field='collectionType', operator='EQUALS', value='Botany')
qs <- QuerySpec$new(conditions=list(qc))

# get the number of specimens
sc$count(qs)$content

```

Note that the count of matches for a given query is also returned by the `query` function. However, `count` is more 
lightweight as it returns an `integer` instead of a `ResultSet` containing `Specimen` objects. 

#### exists 
Check if a record exists, based on its *unitID*:

```{r}
# use SpecimenClient instantiated above
res <- sc$exists('ZMA.INS.1255440')                                        

# content is boolean
res$content
```

#### find 
Return a single specimen given its identifier (Note: the identifier of a specimen is different 
from the unitID, see also [here](http://docs.biodiversitydata.nl/en/latest/doc-spec-services/specimen/#specids)):

```{r}
id <- "RMNH.MAM.17209.B@CRS"
res <- sc$find(id)

# content is single specimen object
res$content
```

#### find_by_ids
Same as `find`, but takes multiple IDs:

```{r}
ids <- "RMNH.INS.657083@CRS,L.1589244@BRAHMS"
res <- sc$find_by_ids(ids)
```

#### find_by_unit_id
Find a specimen by its `unitID`:
```{r}
unitID <- "RMNH.MAM.1513"
res <- sc$find_by_unit_id(unitID)
```

### Aggregation services
Aggregation services group available data according to different criteria. 

#### get_distinct_values
This method takes a specific field as an argument and returns all possible values and the frequency for that 
field in the data. Below we get all possible values for the country in which a specimen was collected. 

```{r results='hide'}
sc$get_distinct_values("gatheringEvent.country")
```

**Note**: By default, `get_distinct_values` lists only the first 10 hits. The above query thus does not reflect the 
distinct values in the hole dataset.
This number can be increased with e.g. setting the `size` parameter in a `QuerySpec` object passed to the method.

```{r results='hide'}
sc$get_distinct_values("gatheringEvent.country",
                       querySpec = QuerySpec$new(size = 10000))
```	

#### count_distinct_values
Instead of returning all different values for a given field, this method does a mere count:

```{r results='hide'}
sc$count_distinct_values("gatheringEvent.country")
```

#### get_distinct_values_per_group
Suppose you want to add another filter to the above query of retrieving all distinct values 
for a given field, such as splitting the results by their respective source system:

```{r results='hide'}
sc$get_distinct_values_per_group("sourceSystem.name", "gatheringEvent.country")
```

#### count_distinct_values_per_group
If no return of the actual values is needed, a simple count per group is done as follows:

```{r results='hide'}
sc$count_distinct_values_per_group("sourceSystem.name",
                                   "gatheringEvent.country")$content
```

### Metadata services

Specimen Metadata services include the same standard metadata services as for the
other data types:

```{r}
# get all paths for the Specimen datatype 
sc$get_paths()$content

# get info e.g. for field collectionType
sc$get_field_info()$content$collectionType

# get all settings
sc$get_settings()

# get specific setting
sc$get_setting("index.max_result_window")$content

# check if operator is allowed
sc$is_operator_allowed("gatheringEvent.continent", "STARTS_WITH")$content
```


#### Available paths and fields

All fields can be retrieved with `get_paths` and specific information 
on the fields, such as allowed operators etc. with `get_field_info`

```{r results='hide'}
sc$get_paths()$content
sc$get_field_info()$content
```

#### Settings
Multimedia-specific settings can be retrieved with `get_settings` and a specific setting
with `get_setting`:

```{r}
sc$get_settings()$content
sc$get_setting("index.max_result_window")$content
```

#### Operators
To test if a certain operator can be used for a multimedia query:

```{r}
sc$is_operator_allowed("identifications.defaultClassification.genus",
                       "STARTS_WITH")$content
```

### DwCA download services
In addition to query services that return JSON formatted data, the NBA also offers
the export of [Darwin Core Archive (DwCA)](https://en.wikipedia.org/wiki/Darwin_Core_Archive) files.
These files are by default zip archives, please refer to 
our [official API documentation](https://en.wikipedia.org/wiki/Darwin_Core_Archive)
for more information. 

#### Static download 
Static download services offer the download of predefined datasets.
The sets that are available for download can be queried with 
`dwca_get_data_set_names`:

```{r}
sc$dwca_get_data_set_names()$content
```

A dataset can then be downloaded using `dwca_get_data_set`. A filename can be given as argument, 
if none is given, the DwCA archive is written to *download-YYYY-MM-DDThh:mm.zip* in the current working directory.

```{r results='hide'}
# download dataset 'porifera' to temporary file
filename <- tempfile(fileext=".zip")
sc$dwca_get_data_set('porifera', filename=filename)                                                                               
```

#### Dynamic download
The dynamic download function `dwca_query` allows for download of arbitrary sets, defined by the user's query.
The arguments to this methods are similar to `query`, plus the filename:

```{r results='hide'}
# download all specimen of genus 'Hydrochoerus'
filename <- tempfile(fileext = ".zip")
qs <-
  QuerySpec$new(conditions = list(
    QueryCondition$new(
      field = "identifications.defaultClassification.genus",
      operator = "EQUALS",
      value = "Hydrochoerus"
    )
  ))
sc$dwca_query(querySpec = qs, filename = filename)
```

## Taxonomic data services {.tabset}

### Query Services

#### query
Query for taxon document with given search criteria. Example:

```{r results='hide'}
# query for taxa of genus 'Sedum' that are in the Netherlands Soortenregister (NSR)
tc <- TaxonClient$new()
qc <-
  QueryCondition$new(field = "acceptedName.genusOrMonomial",
                     operator = "EQUALS",
                     value = "Sedum")
qc2 <-
  QueryCondition$new(field = "sourceSystem.code",
                     operator = "EQUALS",
                     value = "NSR")
qs <- QuerySpec$new(conditions = list(qc, qc2))
tc$query(qs)
```

#### download_query
The `query` function is limited to retrieve 50000 taxa at once (this is determined
in the parameter `index.max_result_window`, the value is retrievable using the `getSettings` method
in the metadata section). In order to provide access for a larger amount of data, 
the `query_download` takes the same arguments as `query`, but download the data as a gzip stream under the hood. 
Unlike `query`, `query_download` returns a list of taxon objects instead of a `ResultSet`.

### Data access services

#### count
For a given query, do not return `Taxon` objects but the mere count. Example

```{r}
# get counts for taxa of genus 'Sedum' that are in the
#  Netherlands Soortenregister (NSR)
qc <-
  QueryCondition$new(field = "acceptedName.genusOrMonomial",
                     operator = "EQUALS",
                     value = "Sedum")
qc2 <-
  QueryCondition$new(field = "sourceSystem.code",
                     operator = "EQUALS",
                     value = "NSR")
qs <- QuerySpec$new(conditions = list(qc, qc2))
tc$count(qs)
```

#### find
Returns a taxon object given its identifier:

```{r}
tc$find("27706109@COL")$content
```

#### find_by_ids
Given a string with comma-separated identifiers, returns a list of taxon objects:

```{r}
ids <- "27706109@COL,27704140@COL,27706110@COL,27706111@COL,27706108@COL"
res <- tc$find_by_ids(ids)
```		

### Aggregation services

#### get_distinct_values
This method takes a specific field as an argument and returns all possible values and the frequency for that 
field in the data. Example: get all data source systems for taxon objects: 

```{r}
tc$get_distinct_values("sourceSystem.name")$content
```

### Metadata services

Taxon Metadata services include the same standard metadata services as for the
other data types:

```{r}
# get all paths for the Taxon datatype 
tc$get_paths()$content

# get info e.g. for field collectionType
tc$get_field_info()$content$synonyms.author

# get all settings
tc$get_settings()

# get specific setting
tc$get_setting("index.max_result_window")$content

# check if operator is allowed
tc$is_operator_allowed("synonyms.author", "EQUALS")$content
```

#### Available paths and fields

All fields can be retrieved with `get_paths` and specific information 
on the fields, such as allowed operators etc. with `get_field_info`

```{r results='hide'}
tc$get_paths()$content
tc$get_field_info()$content
```

#### Settings
Multimedia-specific settings can be retrieved with `get_settings` and a specific setting
with `get_setting`:

```{r}
tc$get_settings()$content
tc$get_setting("index.max_result_window")$content
```

#### Operators
To test if a certain operator can be used for a mutimedia query:

```{r}
tc$is_operator_allowed("identifications.defaultClassification.genus",
                       "STARTS_WITH")$content
```

### DwCA download services
The taxonomic information in the NBA is also available as Darwin-Core archive files.

#### Static download 
Static download services offer the download of predefined datasets.
The sets that are available for download can be queried with 
`dwca_get_data_set_names`:

```{r}
tc$dwca_get_data_set_names()$content
```

A dataset can then be downloaded using `dwca_get_data_set`. A filename can be given as argument, 
if none is given, the DwCA archive is written to *download-YYYY-MM-DDThh:mm.zip* in the current working directory.

```{r results='hide'}
# download dataset 'nsr' to temporary file
filename <- tempfile(fileext=".zip")
tc$dwca_get_data_set('nsr', filename=filename)                                                                               
```

#### Dynamic download
The dynamic download function `dwca_query` allows for download of arbitrary sets, defined by the user's query.
The arguments to this methods are similar to `query`, plus the filename:

```{r results='hide'}
# download all taxa for genus 'Clematis'
filename <- tempfile(fileext = ".zip")
qs <-
  QuerySpec$new(conditions = list(
    QueryCondition$new(
      field = "defaultClassification.genus",
      operator = "EQUALS",
      value = "Clematis"
    )
  ))
tc$dwca_query(querySpec = qs, filename = filename)
```


## Geographic data services {.tabset}

### Query Services

The GeoArea query service allows for detailed search within the 
fields of a GeoArea object. As for the other data types, query parameters
can be either given as a `list` or as a `QuerySpec` object.
Below, we make a simple query to get the `GeoArea` object for the Netherlands:

```{r}
# instantiate client for geo areas
gc <- GeoClient$new()

# query for GeoArea of the Netherlands
qc <-
  QueryCondition$new(field = "locality",
                     operator = "EQUALS",
                     value = "Netherlands")
qs <- QuerySpec$new(conditions = list(qc))
res <- gc$query(qs)

# get item
res$content$resultSet[[1]]$item
```

### Data access services

#### get_geo_json_for_locality
This is a convenience function to directly extract the a GeoJSON object for a 
specific locality. [GeoJSON](https://en.wikipedia.org/wiki/GeoJSON) is a 
popular format for storing geographical point- and polygon data. 
To e.g. extract the GeoJSON polygon representation for the Netherlands:

```{r results='hide'}
loc <- "Netherlands"
res <- gc$get_geo_json_for_locality(loc)
```

Results are returned as a `list` by default, but can be easily converted to 
a JSON string, e.g.

```{r results='hide'}
jsonlite::toJSON(res$content)
```

#### count
For any given query (with `QuerySpec` or not), returns the count of matches instead of `GeoArea` objects:

```{r}
# return count of all GeoAreas
res <- gc$count()
res$content
```

### Aggregation services

#### get_distinct_values
This function returns all values present for a certain field, and their counts:

```{r}
gc$get_distinct_values("areaType")$content
```

### Metadata services
Geo Metadata services include the same standard metadata services as for the
other data types:

```{r}
# get all paths for the GeoArea datatype 
gc$get_paths()$content

# get info e.g. for field 'areaType'
gc$get_field_info()$content$areaType

# get all settings
gc$get_settings()

# get specific setting
gc$get_setting("index.max_result_window")$content

# check if operator is allowed
gc$is_operator_allowed("locality", "STARTS_WITH")$content
```

## Multimedia services {.tabset}

Multimedia services are accessible with a `MultimediaClient`, instantiated
as follows:

```{r}
mc <- MultimediaClient$new()
```

### Query Services

As for the other data types, the `query` method enables simple and complex queries
using a list or a `QuerySpec` object to specify query parameters. 

```{r}
# example of multimedia query passing parameters as a list
mc$query(queryParams = list(collectionType = 'Cnidaria'))$content

# example of multimedia query using QuerySpec: get the first 100
# multimedia items associated with a specimen with name starting with "Ba"
qc <-
  QueryCondition$new(field =
                       "identifications.scientificName.fullScientificName",
                     operator =
                       "STARTS_WITH",
                     value =
                       "Qu")
qs <- QuerySpec$new(conditions = list(qc), size = 100)
res <- mc$query(qs)

# check if scientific names indeed start with 'Qu'
sapply(res$content$resultSet, function(x)
  x$item$identifications[[1]]$scientificName$fullScientificName)
```


### Data access services

#### count
As for the other data types, a `count` function returns counts
instead of the actual objects:

```{r}
# count all multimedia documents
mc$count()$content
```

### Aggregation services

#### get_distinct_values
This function returns all values present for a certain field, and their counts. Example: retrieve
all different licenses and their counts"

```{r}
mc$get_distinct_values("license")$content
```

### Metadata services

#### Available paths and fields

All fields can be retrieved with `get_paths` and specific information 
on the fields, such as allowed operators etc. with `get_field_info`

```{r results='hide'}
mc$get_paths()$content
mc$get_field_info()$content
```

#### Settings
Multimedia-specific settings can be retrieved with `get_settings` and a specific setting
with `get_setting`:

```{r}
mc$get_settings()$content
mc$get_setting("index.max_result_window")$content
```

#### Operators
To test if a certain operator can be used for a mutimedia query:

```{r}
mc$is_operator_allowed("identifications.scientificName.fullScientificName",
                       "STARTS_WITH")$content
```

## Metadata services {.tabset}

Metadata services provide miscellaneous information about the data available via the NBA. Note that 
there is also type-specific metadata for each data type (e.g. `Specimen`) which can be retrieved 
with the specific client of that class. Here we show the available methods for
the `MetadataClient` which gives general, non-type specific metadata.
The client is instantiated in the standard way:

```{r}
mc <- MetadataClient$new()
```


#### Controlled vocabularies
The vocabularies for some fields are controlled by 
dictionaries with allowed values. For the sex of a museum specimen, for instance, only the
terms *male*, *female*, *mixed* and *hermaphrodite* are allowed to be assigned to the specimen.
The fields for which controlled lists are available can be retrieved as follows:

```{r}
mc$get_controlled_lists()$content
```

and for each field that has a controlled vocabulary, there is a separate function to retrieve the
allowed values:

```{r}
mc$get_controlled_list_taxonomic_status()$content
mc$get_controlled_list_specimen_type_status()$content
mc$get_controlled_list_sex()$content
mc$get_controlled_list_phase_or_stage()$content
```

#### Miscellaneous

##### Dates
To maintain data integrity, dates have to be coded in specific formats
in our systems. For instance, *yyyy-MM-dd* is a valid format.
Allowed formats can be retrieved as follows:

```{r}
mc$get_allowed_date_formats()$content
```

##### Services list
The method `get_rest_services` returns a list of all services
in the NBA as objects of type `RestService`.

```{r}
# get the endPoint of the first rest service in the services list
mc$get_rest_services()$content[[1]]
```

##### Settings
Similar to the document-specific metadata services, we can get general settings with 
the `MetaDataClient`:

```{r}
#get all settings
mc$get_settings()$content

# get value for specific setting
mc$get_setting("operator.contains.min_term_length")$content
```
