---
title: "The oldest tomato specimen in the world"
date: "`r format(Sys.time(), '%b %d, %Y')`"
output:
  html_document:
    highlight: tango
    code_folding: show
    theme: sandstone
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
---

# The 'En Tibi' tomato specimen

## Using the SpecimenClient

For the different data types in the NBA (Specimen, Taxon, Multimedia, Geo),
there are respective client classes (`SpecimenClient`, `TaxonClient`,
`MultimediaClient`, `GeoClient`). To retreive electronic specimen records,
we instantiate a `SpecimenClient`:

```{r}
library(nbaR)
sc <- SpecimenClient$new()
```

To get an overview of what we can do with such a client, `?SpecimenClient` will
give us some description and list all the methods that are available for this client.
To get an overview of what nbaR stores in a `Specimen` object, we could use the function
`get_paths`, which lists everything that can be queried for a specimen:

```{r}
res <- sc$get_paths()
```

Note that client class methods (and other class members) are accessed with a `$`.
The response from the taxon client is stored in `res`, an object of class `Response`.
Have a look at `?Response` to see what this object contains. But how to get the data?
It is in the field `content`.

```{r}
res$content
```

Suppose that we do not know the scientific name of the tomato plant,
there are two fields that store a vernacular name for a specimen:

* `identifications.vernacularNames.name`
* `identifications.taxonomicEnrichments.vernacularNames.name`

The first one is the vernacular name that is exactly equal to the one in the
*source database*. Since often, a vernacular name is missing in the source database,
a *taxonomic enrichment* is performed during the import of the data from the source database
in which, using among others the *Catalogue of life*, species names are enriched with synonyms and
vernacular names.

## Simple queries

Let's try the first field and query for all `Specimen` records with `identifications.vernacularNames.name` field
matching the name `tomato`, we can use the client's `query` function:

```{r}
res <-
  sc$query(queryParams = list(identifications.vernacularNames.name = "tomato"))
```

Note that the `queryParams` arguments takes a list, so
we could search for more criteria. How many hits do we got?

```{r}
res$content$totalSize
```

No hits, thus no record has the vernacular name equal to `tomato`. We therefore
search in `identifications.taxonomicEnrichments.vernacularNames.name`:

```{r}
res <-
  sc$query(
    queryParams =
      list(
        identifications.taxonomicEnrichments.vernacularNames.name =
          "tomato"
      )
  )
# how many hits?
res$content$totalSize
```

Again, no hits.

## More complex queries

Let's try partial matching. For this, we have to make a
more complicated query, involving `QuerySpec` and `QueryCondition` objects,
which are very powerful for constructing complex and nested queries.

A `QueryCondition` can be specified as follows:

```{r}
qc <-
  QueryCondition$new(
    field = "identifications.taxonomicEnrichments.vernacularNames.name",
    operator = "MATCHES",
    value = "tomato"
  )
```

Note that we use the operator `MATCHES` that does a partial instead of an exact match
(which would be `EQUALS`).

From one or multiple `QueryConditions`, a `QuerySpec` can be created, and used as input
for he `query` function.

```{r}
qs <- QuerySpec$new(conditions = list(qc))

# do the query with SpecimenClient
res <- sc$query(querySpec = qs)

# how many hits?
res$content$totalSize
```

Finally, we have some hits. The content of the response of the `query` function
is always of type `QueryResult` (see also `?QueryResult`) which has the fields
`totalSize` (as used above) and a `resultSet`, a list which stores the actual data.
From the `resultSet`, a single `Specimen` object can be obtained as follows:

```{r}
# retreive the first Specimen object
sp <- res$content$resultSet[[1]]$item

# check if it is really a specimen
class(sp)

# list fields and methods of the object
sp
```

**Note:** By default, the NBA (and thus nbaR) returns the first 10 hits.
To fetch more records, we need to specify this in the `QuerySpec` object.


```{r}
# specify size in QuerySpec obejct
qs <- QuerySpec$new(conditions = list(qc), size = 1000)

# perform query
res <- sc$query(querySpec = qs)

# do we have all records?
res$content$totalSize == length(res$content$resultSet)
```

## Sorting

We now want to see which one is the oldest specimen. We could, of course
do the sorting in R
(e.g. `sort(unlist(lapply(res$content$resultSet, function(x)x$item$gatheringEvent$dateTimeBegin)))`),
but there is also functionality to do this directly in the query.
This can be done with objects of type `?SortField`, which can take a
path specifying on what to sort, and whether to sort ascending or
descending. We will sort our results by the values of the field
`gatheringEvent.dateTimeBegin`:


```{r}
# specify field to sort
sf <- SortField$new(path = "gatheringEvent.dateTimeBegin", sortOrder = "asc")

# make querySpec using sortField
qs <- QuerySpec$new(conditions = list(qc), size = 1000, sortFields = list(sf))

# do the query
res <- sc$query(querySpec = qs)
```

Now, the first result should be the specimen with a gathering event the furthest
in the past.

```{r}
sp <- res$content$resultSet[[1]]$item

# get date
sp$gatheringEvent$dateTimeBegin
```

## Multimedia content

Many specimens have associated multimedia content, such as photos or videos.
To retrieve e.g. the URL of the first multimedia item of our specimen:

```{r}
# get multimedia URL
url <- sp$associatedMultiMediaUris[[1]]$accessUri
url

# display image
library(knitr)
include_graphics(url)
```

# Retrieving specimen records with lat/long coordinates

Geo-referenced specimens are an invaluable resource for biogeographic analyses.
The below example shows how to extract the tomato specimen (species *Solanum lycopersicum*)
that are geo referenced. The coordinates, stored in the field *gatheringEvent.siteCoordinates*
are then extracted and the locations are plotted on a world map.

**Note:** When querying for records with non-empty values for a specific field, 
we use the operator `NOT_EQUALS` in combination with the value `NULL`. 

```{r}		
# conditions for specimens for Solanum lycopersicum
qc <-
  QueryCondition$new(
    field =
      "identifications.defaultClassification.genus",
    operator =
      "EQUALS",
    value =
      "Solanum"
  )
qc2 <-
  QueryCondition$new(
    field =
      "identifications.defaultClassification.specificEpithet",
    operator =
      "EQUALS",
    value = "lycopersicum"
  )

# coditions for lat/long coordinates to be present
qc3 <-
  QueryCondition$new(
    field =
      "gatheringEvent.siteCoordinates.longitudeDecimal",
    operator =
      "NOT_EQUALS",
    value =
      NULL
  )
qc4 <-
  QueryCondition$new(
    field =
      "gatheringEvent.siteCoordinates.latitudeDecimal",
    operator =
      "NOT_EQUALS",
    value =
      NULL
  )
qs <-
  QuerySpec$new(conditions = list(qc, qc2, qc3, qc4), size = 1000)

# do query
res <- sc$query(querySpec = qs)

# extract coordinates
lat <-
  sapply(res$content$resultSet, function(x) {
    x$item$gatheringEvent$siteCoordinates[[1]]$latitudeDecimal
  })
long <-
  sapply(res$content$resultSet, function(x) {
    x$item$gatheringEvent$siteCoordinates[[1]]$longitudeDecimal
  })

# plot on world map
library("maps")
map(
  "world",
  fill = TRUE,
  col = "white",
  bg = "lightblue",
  ylim = c(-60, 90),
  mar = c(0, 0, 0, 0)
)
points(long, lat, col = "red", pch = 16)
```


